---
title: "Analyzing Forest Fire Data"
author: "RZ"
date: "6/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import the library and the setting working directory
```{r}
library(tidyverse)
library(ggplot2)
setwd("~/Downloads/Dataquest/R")
fire <- read.csv('forestfires.csv')
print(colnames(fire))
```
A single record indicates one fire record

Check the unique values of day and month
```{r}
print(fire %>% pull(month) %>% unique())
fire %>% pull(day) %>% unique()
```

Convert the month and day to categorical variable
```{r}
month_order <- c("jan", "feb", "mar",
                 "apr", "may", "jun",
                 "jul", "aug", "sep",
                 "oct", "nov", "dec")
dow_order <- c("sun", "mon", "tue", "wed", "thu", "fri", "sat")
fire <- fire %>% mutate(
  month = factor(month, levels = month_order),
  day = factor(day, levels = dow_order)
)
```

Check the most happen in which month and day
```{r}
fire_month <- fire %>% group_by(month) %>% summarise(
  count_m = n()
)

fire_day <- fire %>% group_by(day) %>% summarise(
  count_d = n()
)
```

Visualize the counts by month and day
```{r}
fire_month %>% ggplot(aes(x=month, y=count_m)) + geom_col() + labs(
  title = "Number of forest fires in data by month of the year",
  y = "Fire count",
  x = "Month of the Year"
  )

fire_day %>% ggplot(aes(x=day, y=count_d)) + geom_col() + labs(
  title = "Number of forest fires in data by day of the week",
  y = "Fire count",
  x = "Day of the week"
  )
```

Check each variable relationship to month
```{r}
fire_long <- fire %>% 
  pivot_longer(
    cols = c("FFMC", "DMC", "DC", 
             "ISI", "temp", "RH", 
             "wind", "rain"),
    names_to = "data_col",
    values_to = "value"
  )

fire_long %>% ggplot(aes(x=month, y=value)) +geom_boxplot() + facet_wrap(vars(data_col), scale = 'free_y') + labs(
  title = "Variable changes over month",
  x = "Month",
  y = "Variable value")
```

Check the relationship of the areas and each variable
```{r}
fire_long %>% ggplot(aes(x=value, y=area)) +geom_point() + facet_wrap(vars(data_col), scale = 'free_x') + labs(
  title = "area changes over variables",
  x = "Variable value",
  y = "area")
```

remove the data with area smaller than 300 to remove outliers
```{r}
fire_long %>% filter(area <300) %>% ggplot(aes(x=value, y=area)) +geom_point() + facet_wrap(vars(data_col), scale = 'free_x') + labs(
  title = "area changes over variables",
  x = "Variable value",
  y = "area")
```
