---
title: "NYC Schools Perceptions"
author: "RZ"
date: "6/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import the necessary files and packages
```{r}
library(tidyverse)
library(readr)
library(dplyr)
library(purrr)
library(ggplot2)
library(tidyr)
# All data are from 2011 NYC online data
combined <- read_csv('combined.csv')
survey <- read_tsv('2011 data files online/masterfile11_gened_final.txt')
survey_75 <- read_tsv('2011 data files online/masterfile11_d75_final.txt')

```

Filter the data frame to remove unnecessary columns
```{r}
survey_select <- survey %>% select(dbn:aca_tot_11) %>% filter(schooltype=='High School')
survey_75_select <- survey_75 %>% select(dbn:aca_tot_11)
```

Combine the survey data with selected conditions above
```{r}
survey_total <- bind_rows(survey_select, survey_75_select) 
survey_total <- survey_total %>% rename(DBN=dbn)
combined_survey <- combined %>% left_join(survey_total, by='DBN')
```

Find the correlation and visualize the correlation
```{r}
# Find the correlation matrix
cor_mat <- combined_survey %>% select(avg_sat_score, saf_p_11:aca_tot_11) %>% cor(use = "pairwise.complete.obs")
# Covert the correlation matrix to tibble which as variable as row names
cor_tib <- cor_mat %>%as_tibble(rownames = "variable")
# Find strong correlations
strong_cors <- cor_tib %>%select(variable, avg_sat_score) %>%filter(avg_sat_score > 0.25 | avg_sat_score < -0.25)  
```

Visualize the avg_sat_score to other strong correlation variables
```{r}
create_scatter <- function(x, y) {ggplot(data = combined_survey) + aes_string(x = x, y = y) + geom_point(alpha = 0.3) + theme(panel.background = element_rect(fill = "white"))}

x_var <- strong_cors$variable[2:5]
y_var <- "avg_sat_score"
  
map2(x_var, y_var, create_scatter)
```

Reshape the data so that you can investigate differences in student, parent, and teacher responses to survey questions.
```{r}
combined_survey_gather <- combined_survey %>%
  pivot_longer(cols = saf_p_11:aca_tot_11,
               names_to = "survey_question",
               values_to = "score")
# Extract values from the string
combined_survey_gather <- combined_survey_gather %>%
  mutate(response_type = str_sub(survey_question, 4, 6)) %>%   
  mutate(question = str_sub(survey_question, 1, 3))

combined_survey_gather <- combined_survey_gather %>%
  mutate(response_type = ifelse(response_type  == "_p_", "parent", 
                                ifelse(response_type == "_t_", "teacher",
                                       ifelse(response_type == "_s_", "student", 
                                              ifelse(response_type == "_to", "total", "NA")))))

# Make a box plot to see if there appear to be differences in how the three groups of responds (parents, students, and teachers) answered the four questions. 
combined_survey_gather %>%
  filter(response_type != "total") %>%
  ggplot(aes(x = question, y = score, fill = response_type)) +
  geom_boxplot()
```

